---
title: "Practical machine learning  lectures 4"
author: "Kirill Setdekov"
date: "15 01 2020"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	cache = TRUE
)
```

## Task:
* The goal of your project is to predict the manner in which they did the exercise. 
* This is the "classe" variable in the training set. You may use any of the other variables to predict with. 
* You should create a report describing how you built your model, 
* how you used cross validation, 
* what you think the expected out of sample error is, and 
* why you made the choices you did. 
* You will also use your prediction model to predict 20 different test cases.

```{r}
if(!file.exists("./data")) {
     dir.create("./data")
}
if (!file.exists("./data/training.csv") |
    !file.exists("./data/testing.csv")) {
     fileUrl1 <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
     fileUrl2 <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
     download.file(fileUrl1, destfile = "./data/training.csv")
     download.file(fileUrl2, destfile = "./data/testing.csv")
}

buildData <- read.csv("./data/training.csv")
validation <- read.csv("./data/testing.csv")

head(buildData, 2)
head(validation, 2)
```


```{r wageex}
library(caret)
allmissing <- sapply(validation, function(x)!all(is.na(x)))
buildData[is.na(buildData)] <- 0
validation[is.na(validation)] <- 0
buildData <- buildData[,allmissing]
validation <- validation[,allmissing]

buildData <- buildData[,-1] 
validation <- validation[,-1] 

inTrain <- createDataPartition(y=buildData$classe, p=0.7, list = FALSE)
training <- buildData[inTrain,]
testing <- buildData[-inTrain,]





```

### build 2 models

```{r warning=FALSE}
library(party)

# mod1 <- train(classe~., method ="glm", data = training)
mod2 <- train(classe~., method ="rf", data = training, trControl = trainControl(method = "cv"),number = 3)
mod3 <- train(classe~., method ="rpart", data = training)
# mod4 <- train(classe~., method ="ctree", data = training)
# mod5 <- mob(classe~age+education|year+maritl+race+region+jobclass+health+health_ins,data = training)

```
### compare them
```{r dependson=c(-1)}
# pred1 <- predict(mod1, testing)
pred2 <- predict(mod2, testing)
pred3 <- predict(mod3, testing)
# pred4 <- predict(mod4, testing)
# pred5 <- predict(mod5, testing)
                 
qplot(pred1,pred2, colour = wage, data = testing)
qplot(pred3,pred4, colour = wage, data = testing)
table(pred3, testing$classe)
table(pred2, testing$classe)

library(rattle)
fancyRpartPlot(mod3$finalModel)

plot(mod4$finalModel)
plot(mod5)

qplot(wage,pred5, colour = wage, data = testing)

```

### make a model that combines predictors

```{r dependson=c(-1, -2)}
predDF2 <- data.frame(pred1, pred2, wage = testing$wage)
#variant with 5
predDF5 <- data.frame(pred1, pred2, pred3, pred4, pred5, wage = testing$wage)

combModFit2 <- train(wage~., method = "gam", data = predDF2)
combModFit5 <- train(wage~., method = "gam", data = predDF5)
combPred2 <- predict(combModFit2,predDF2)
combPred5 <- predict(combModFit5,predDF2)

```

#### testing errors
```{r}
MLmetrics::RMSE(y_pred = pred1,y_true = testing$wage)
MLmetrics::RMSE(y_pred = pred2,y_true = testing$wage)
MLmetrics::RMSE(y_pred = combPred2,y_true = testing$wage)

MLmetrics::RMSE(y_pred = pred3,y_true = testing$wage)
MLmetrics::RMSE(y_pred = pred4,y_true = testing$wage)
MLmetrics::RMSE(y_pred = pred5,y_true = testing$wage)
MLmetrics::RMSE(y_pred = combPred5,y_true = testing$wage)

plot(combModFit5$finalModel)
```
### on a validation
```{r}
pred1V <- predict(mod1, validation)
pred2V <- predict(mod2, validation)
predVDF2 <- data.frame(pred1 = pred1V, pred2 = pred2V)
combPredV2 <- predict(combModFit2, predVDF2)

pred3V <- predict(mod3, validation)
pred4V <- predict(mod4, validation)
pred5V <- predict(mod5, validation)
predVDF5 <-
    data.frame(
        pred1 = pred1V,
        pred2 = pred2V,
        pred3 = pred3V,
        pred4 = pred4V,
        pred5 = pred5V
    )
combPredV5 <- predict(combModFit5,predVDF5)
MLmetrics::RMSE(y_pred = pred1V,y_true = validation$wage)
MLmetrics::RMSE(y_pred = pred2V,y_true = validation$wage)
MLmetrics::RMSE(y_pred = combPredV2,y_true = validation$wage)
## 5 part model
MLmetrics::RMSE(y_pred = pred3V,y_true = validation$wage)
MLmetrics::RMSE(y_pred = pred4V,y_true = validation$wage)
MLmetrics::RMSE(y_pred = pred5V,y_true = validation$wage)
MLmetrics::RMSE(y_pred = combPredV5,y_true = validation$wage)
```


